---
title: 'Solutions chapter 1 - Geocomputation with R'
author: Jorge Cimentada
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r knitr-setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE,
                      fig.path = "../figs/",
                      fig.align = "center",
                      fig.asp = 0.618,
                      out.width = "80%")
```


```{r }
library(leaflet)
popup = c("Robin", "Jakub", "Jannes")
leaflet() %>%
  addProviderTiles("NASAGIBS.ViirsEarthAtNight2012") %>%
  addMarkers(lng = c(-3, 23, 11),
             lat = c(52, 53, 49), 
             popup = popup)
```


1. Think about the terms ‘GIS’, ‘GDS’ and ‘geocomputation’ described above. Which (if any) best describes the work you would like to do using geo* methods and software and why?

I'm more interested in doing geocomputation, which in general related to solving spatial problems (both for visualization and for spatial statistics) both for pure scientific purposes as well as for day-to-day problems. This encompasses both some type of Geographic Data Science but also on the deep understanding of micro spatial patterns.

2. Provide three reasons for using a scriptable language such as R for geocomputation instead of using an established GIS program such as QGIS.

- Reproducibility
- General purpose skills
- Flexibility

3. Name two advantages and two disadvantages of using mature vs recent packages for geographic data analysis (for example sp vs sf).

* Using recent packages

- Advantages:
  - Up to date algorithms
  - Usually, more intuitive and easy user interface
  
- Disadvantage
  - More prone to bugs
  - More unstable on the long terms (breaking changes)
